Automated verification of stand-alone solar photovoltaic systems

### 2018-11-15 Tex/LaTeX2e Style File/Trindade_IEEE_journal_r2.tex
paper update

%--------------------------------------------------
\section{Automated verification}
\label{sec:AutomatedVerification}
%--------------------------------------------------

%It is necessary to keep in mind that 
Validation is the process of determining whether a design meets the needs of the user, whereas verification is the process of determining whether a design meets a set of requirements, specifications, and regulations~\cite{ClarkeHV18}.

If the requirements, specifications, and regulations are given in a formal language, then it may be possible to automate verification, resulting in a process known as formal verification. Verification may form part of a validation process.
%, but in general, validation cannot be formalized because it relates a system design to intent.

%Simulation may also be used for validation, but it is more problematic for verification.
In order to use simulation for verification, it is necessary to ensure adequate coverage of operating conditions, scenarios, and system inputs. Testing can also be used for validation, but for the same reasons, it is too problematic for verification \textcolor{red}{Why is it problematic?!?}.

Verification procedure is an intelligent \textcolor{red}{What do you mean by intelligent?} exhaustive search of the state space of the design \cite{ClarkeHV18}. In addition, formal verification is a systematic approach that applies mathematical reasoning to obtain guarantees about the correctness of a system \cite{Forejt2011}. One successful method in this domain is model checking \textcolor{red}{please cite here the reference given above}.
 % 
%Model checking is an automatic verification technique, as defined by \cite{ClarkeHV18}. Model checking was originally developed reasoned about finite state of concurrent systems, nowadays is mainly used to hardware and software verification, but can be applied to any kind of system.

@@

The model checking algorithm can be described as \textcolor{red}{add citation here}:

\begin{itemize}
\item Given the model $ 'M' $ and a $CTL$ formula $ \phi $ as input;  
\item Model checking algorithm provides all the states of model $ M $ which satisfies $ \phi $;  
\item It returns $YES$ if $ \phi $ is $TRUE$, or returns $NO$ if $ \phi $ is $FALSE$.

\end{itemize}
Specifically for the $FALSE$ verification result, the algorithm returns a counterexample that is useful as diagnostic of the system, to discover in which situation the model is violated. That is the most important advantage of the use of model checking \cite{ClarkeHV18}. 
%Among the other advantages can be listed: there is no need of proofs (the algorithm is not a deductive procedure), there is no problem with partial specifications of the system, logics can easily express many concurrency properties, is fast (compared to other rigorous methods such as interactive theorem proving). 
However, there is a main disadvantage of model checking: the state explosion problem. 
%
%The model checking problem can be defined as \cite{ClarkeHV18}: 
%
%\begin{itemize}
%\item Let $M$ be a Kripke structure (i.e., state transition graph);
%\item $f$ be the specification in temporal logic (a formula);
%\item Find all states $s$ of $M$ such that $M , s \models f$
%\end{itemize}

%Fig. \ref{fig:modelcheckstruc} shows the structure of a 
In a typical model checking system, a preprocessor extracts a state transition graph from a system (program or circuit).
%
%\begin{figure}[h]
%\includegraphics[width=0.4\textwidth]{modelcheckstruc}
%\centering
%\caption{Model Checker structure (Source: \cite{ClarkeHV18})}
%\label{fig:modelcheckstruc}
%\end{figure}
%
%Here it is worth to mention that the term "model" is not the meaning taken from the dictionary. The problem is not dealing with an abstraction of the actual system under study. 
%
Fig.~\ref{fig:systemverif} shows the process to convert a real system to a model to be verified by a model checking procedure \textcolor{red}{Please try to adapt this figure to our context. How can it be applied to verify PV?}.

\begin{figure}[h]
\includegraphics[width=0.4\textwidth]{systemverif}
\centering
\caption{From real system verification to model checking (adapted from \cite{ClarkeHV18}).}
\label{fig:systemverif}
\end{figure}

In order to solve the problem of state explosion, many different techniques were developed at the last decades \textcolor{red}{Please cite them here!!!}. One of the promising is the Bounded Model Checking (BMC). BMC is a method that checks the model up to a given path in the path length. BMC algorithms traverses a finite state machine for a fixed number of steps, $ k $, and checks whether violation occurs with this bound. It uses Boolean Satisfiability (SAT) solvers, where SAT means satisfiability. SAT problem is a problem of determining if there are certain conditions or interpretation that satisfy a given Boolean expression \cite{ClarkeHV18}. %SAT solvers are used in BMC, such that i
%If there are some Boolean function, t
The solver would search the model for conditions (value of variables) that would make the formula $TRUE$. If a SAT Solver find a substitution for the formula/function then the substitute induces a counterexample.

@@

\textcolor{red}{ESBMC does not use CBMC as its main frontend. Please take a look at this paper and cite it: https://ssvlab.github.io/lucasccordeiro/papers/ase2018.pdf. I would also suggest to remove the above sentence.}

ESBMC is a context-bounded model checker for embedded C/C++ software based on Satisfiability Modulo Theories (SMT) solver, which can use CBMC as front-end.
%The use of SMT, instead of Boolean Satisfiability SAT from the original BMC, 
ESBMC comes as an alternative to overcome limitations of the systems modeling, especially considering that the complexity of these is increasing and the SMT method has high level and richer theories than the SAT to represent the models \textcolor{red}{add citation here}.

\subsection{ESBMC }
ESBMC (or Efficient SMT-based Bounded Model Checker) is an open source, permissively licensed (Apache 2), cross platform bounded model checking for C and C++ programs \cite{GadelhaSBMF}. 
%The efficient SMT based model checker, is a software verification tool for C and C++ code bases. 
Fig. \ref{fig:esbmcarch} shows the tool architecture.

\begin{figure}[h]
\includegraphics[width=0.5\textwidth]{ESBMCarch2}
\centering
\caption{ESBMC architecture (adapted from \cite{GadelhaSBMF}). \textcolor{red}{This architecture is obsolete. Please see the paper https://ssvlab.github.io/lucasccordeiro/papers/ase2018.pdf}}
\label{fig:esbmcarch}
\end{figure}

ESBMC has two alternative front-ends to parse the input program \cite{GadelhaSBMF}. % and generate an Abstract Syntax Tree (AST). %One legacy CBMC-based front-end that supports both C and C++ and a new clang-based front-end that currently only supports C. The data types and bit vectors are created in the front-end when parsing the code. 
Regardless of the chosen front-end, the output is an AST (tree representation of the abstract syntactic structure of source code written in a programming language) that will be used by the GOTO converter to generate a GOTO program, which has simplified control flow and is suitable for bounded unwinding. The next step is the symbolic execution, when the GOTO program is executed (unrolling loops up to the bound $ k $) and converted to Static Single Assignments (SSA) form. SSA form enables the simplification of algorithms and reduction of computational complexity. %, where each variable is a target of exactly one assignment statement in the program text. 
%
%During the symbolic execution, ESBMC aggressively tries to simplify the program \cite{Ramalho}; it propagates all constants and solves any assertions that can be statically determined. This is an important step for the verification, since ESBMC can fully verify programs without calling a solver, if the inputs are deterministic. That characteristic is useful at this work to perform the sizing check of the PV system project. 
Following, the SSA expressions are then encoded using the chosen SMT solver supported by ESBMC. At this paper, it was used the Z3 solver \cite{DeMoura}.
%: Boolector (default) \cite{Brummayer}, Z3 \cite{DeMoura}, MathSAT \cite{Cimatti}, Yices \cite{Dutertre} and CVC4 \cite{Barrett}.

@@

%\subsection{Cyber-physical systems x Energy x ESBMC}
%According \cite{UC}, cyber-physical systems (CPS) are integrations of computation, networking, and physical processes. Embedded computers and networks monitor and control the physical processes, with feedback loops where physical processes affect computations and vice-versa. The economic and societal potential of such systems is vastly greater than what has been realized, and major investments are being made worldwide to develop the technology. The technology builds on the discipline of embedded systems (computers and software embedded in devices) whose principle mission is not computation, such as cars, toys, medical devices, and scientific instruments. CPS integrates the dynamics of the physical processes with those of the software and networking, providing abstractions and modeling, design, and analysis techniques for the integrated whole. 
%
%Energy production, distribution, and optimization are all CPS problems, mas mentioned in \cite{UC}. For example, the smart grid combines multiple electric power production plants with a multiplicity of loads using dynamic load balancing and dynamic pricing with demand-response strategies. Smart buildings integrate sensors into control systems for lighting, HVAC (heating, ventilation, and air conditioning), and safety (as fire monitoring and evacuation). 
%
%Therefore, in the context of this proposal, a cyber-physical system, specifically a solar photovoltaic system, will be mathematically modeled. Then a model checking will be performed to do a verification of the project in order to guarantee that the design of a stand-alone PV solution will meet the requirements, specifications and regulations. 
%
%As described in \cite{Alur}, in program verification, it is aimed to check if a program satisfies its logical specification. Contemporary verification tools vary widely in terms of source languages, verification methodology, and the degree of automation, but they all rely on repeatedly invoking an SMT solver. This work tracks the same path. 
%
%----------------------------------------
\section{Methodology }
\label{sec:Methodology}
%----------------------------------------

%
%Fig. \ref{fig:statetransition} shows the state diagram for the proposed tool. 
%
%\begin{figure}[h]
%\includegraphics[width=0.5\textwidth]{statetransition}
%\centering
%\caption{State transition diagram to the proposed automated verification.}
%\label{fig:statetransition}
%\end{figure}
%
%It is worth mentioning that the steps described here are carried out shortly after the design of the solar photovoltaic system, as soon as the equipment and equipment specifications are defined; before buying and deploying them, as a way to ensure that the project will not fail. 
%
The flowchart of the proposed (automated) verification method is illustrated in Fig.~\ref{fig:flowchartgeneral}. 
The PV input data (load power demand, load consumption of energy, the PV system project sizing data, and the manufacturer's data-sheet information), the formulas to check the sizing project, the mathematical model, the limits of the weather non-deterministic variables, are all written as an ANSI-C code \textcolor{red}{add citation here to the standard}.

@@

\textcolor{red}{I didn't understand this sentence... it is completely broken...} Simply put, instead of the ESBMC find the defect of a software or if a hardware does not meet the some requirements, the tool will verify if the requirements of the solar PV project are met. Detailing the condition of violation, i.e., the state of each variable related to the system.

@@

%At both case studies, it was considered 21 years of historic data from Manaus. 
%According with X and Y, t
The annual average temperature ($T$) in Manaus is from 23$^{o}$C to 32$^{o}$C; and irradiance ($G$) varies from 274 W/m$^{2}$ to 852 W/m$^{2}$ when there is sunlight.  
%and irradiance varies from 0 W/m$^{2}$ to 852 W/m$^{2}$ (with minimum of 274 W/m$^{2}$ during the daytime, when there is sunlight). 
Another characteristic of Manaus, is that the PV generation can be despised from 16:00h to the 8:00h of the next day, independent of the season.
%there is PV generation only between 8:00 h and 16:00 h every day, 
%with zero electric energy generation from 18:00 h to 6:00 h of the next day; and with insignificant generation from 6:00 h to 8:00 h, and from 16:00 h to 18:00 h of the same day. 
Therefore, in order to reduce the computational effort,
% caused by the state explosion inherent of the technique, 
the 24h-day was divided, at the algorithm, in 3 parts of 8 h each (8:00h-16:00h, 16:00h-24:00h, and 00:00h-8:00h); and only the 8:00h-16:00h period with PV generation. The time step was adjusted to 1 hour. 
% and the non-deterministic variables $G$ and $T$ are considered during the formal verification of the system, otherwise, during the other two periods, there is no PV generation and just the power consumption from the backup batteries. 
%Within this 8h-period, $G$ and $T$ are automated verified with different values every one hour.
%, and change their value every 1 h according with the algorithm created using the technique.

Algorithm~\ref{alg:verification-algorithm} describes the pseudo code used to perform the automated verification. \textcolor{red}{Can you please briefly explain this algorithm?}

>>>>

%--------------------------------------------------
\section{Automated verification}
\label{sec:AutomatedVerification}
%--------------------------------------------------
%It is necessary to keep in mind that 
Validation is the process of determining whether a design meets the needs of the user, whereas verification is the process of determining whether a design meets a set of requirements, specifications, and regulations~\cite{ClarkeHV18}.

If the requirements, specifications, and regulations are given in a formal language, then it may be possible to automate verification, resulting in a process known as formal verification. Verification may form part of a validation process.
%, but in general, validation cannot be formalized because it relates a system design to intent.  
While simulation and testing explore some of the possible behaviors and scenarios of the system, leaving open the question of whether the unexplored trajectories may contain a flaw, formal verification conducts an exhaustive exploration of all possible behaviors. Thus, when a design is pronounced correct by a formal verification method, it implies that all behaviors have been explored, and the questions of adequate coverage or a missed behavior become irrelevant \cite{Clarke2012}.
%Simulation may also be used for validation, but it is more problematic for verification.
%In order to use simulation for verification, it is necessary to ensure adequate coverage of operating conditions, scenarios, and system inputs. Testing can also be used for validation, however it is not feasible to cover all the aspects of the design space, and some times the real system must be used instead of a computing model.

Formal verification is a systematic approach that applies mathematical reasoning to obtain guarantees about the correctness of a system \cite{Forejt2011}. One successful method in this domain is model checking \cite{Clarke2012}.
 % 
%Model checking is an automatic verification technique, as defined by \cite{ClarkeHV18}. Model checking was originally developed reasoned about finite state of concurrent systems, nowadays is mainly used to hardware and software verification, but can be applied to any kind of system.

@@

The model checking algorithm can be described as \cite{ClarkeHV18}:

\begin{itemize}
\item Given the model $ 'M' $ and a $CTL$ formula $ \phi $ as input;  
\item Model checking algorithm provides all the states of model $ M $ which satisfies $ \phi $;  
\item It returns $YES$ if $ \phi $ is $TRUE$, or returns $NO$ if $ \phi $ is $FALSE$.  
\end{itemize}
Specifically for the $FALSE$ verification result, the algorithm returns a counterexample that is useful as diagnostic of the system, to discover in which situation the model is violated. That is the most important advantage of the use of model checking \cite{ClarkeHV18}. 
%Among the other advantages can be listed: there is no need of proofs (the algorithm is not a deductive procedure), there is no problem with partial specifications of the system, logics can easily express many concurrency properties, is fast (compared to other rigorous methods such as interactive theorem proving). 
%
%The model checking problem can be defined as \cite{ClarkeHV18}: 
%
%\begin{itemize}
%\item Let $M$ be a Kripke structure (i.e., state transition graph);
%\item $f$ be the specification in temporal logic (a formula);
%\item Find all states $s$ of $M$ such that $M , s \models f$
%\end{itemize}

%Fig. \ref{fig:modelcheckstruc} shows the structure of a 
In a typical model checking system, a preprocessor extracts a state transition graph from a system (program or circuit).
%
%\begin{figure}[h]
%\includegraphics[width=0.4\textwidth]{modelcheckstruc}
%\centering
%\caption{Model Checker structure (Source: \cite{ClarkeHV18})}
%\label{fig:modelcheckstruc}
%\end{figure}
%
%Here it is worth to mention that the term "model" is not the meaning taken from the dictionary. The problem is not dealing with an abstraction of the actual system under study. 
%
Fig.~\ref{fig:systemverif} shows the process to convert a real PV system to a model to be verified by a model checking procedure. % \textcolor{red}{Please try to adapt this figure to our context. How can it be applied to verify PV?}.

\begin{figure}[h]
\includegraphics[width=0.5\textwidth]{systemverif2}
\centering
\caption{From real system verification to model checking (adapted from \cite{ClarkeHV18}).}
\label{fig:systemverif}
\end{figure}

However, there is a main disadvantage of model checking: the state explosion problem. In order to solve this problem, many different techniques were developed at the last decades. One of the first major advances was symbolic model checking with binary decision diagrams (BDDs). In this approach, a set of states is represented by a BDD instead of by listing each state individually, which is often exponentially smaller in practice. Another major advance is the partial order reduction, which exploits the independence of actions in a system with asynchronous composition of processes. A third major advance is counterexample-guided abstraction refinement, which adaptively tries to find an appropriate level refinement, not burdened with irrelevant detail that slows down verification \cite{Clarke2012}. However, one of the promising is the Bounded Model Checking (BMC).

BMC is a method that checks the model up to a given path in the path length. BMC algorithms traverses a finite state machine for a fixed number of steps, $ k $, and checks whether violation occurs with this bound. It uses Boolean Satisfiability (SAT) solvers, where SAT means satisfiability. SAT problem is a problem of determining if there are certain conditions or interpretation that satisfy a given Boolean expression \cite{ClarkeHV18}. %SAT solvers are used in BMC, such that i
%If there are some Boolean function, t
The solver would search the model for conditions (value of variables) that would make the formula $TRUE$. If a SAT Solver find a substitution for the formula/function then the substitute induces a counterexample.

@@

%\textcolor{red}{ESBMC does not use CBMC as its main frontend. Please take a look at this paper and cite it: https://ssvlab.github.io/lucasccordeiro/papers/ase2018.pdf. I would also suggest to remove the above sentence.}

ESBMC is a context-bounded model checker for embedded C/C++ software based on Satisfiability Modulo Theories (SMT) solver. %, which can use CBMC as front-end.
%The use of SMT, instead of Boolean Satisfiability SAT from the original BMC, 
ESBMC comes as an alternative to overcome limitations of the systems modeling, especially considering that the complexity of these is increasing and the SMT method has high level and richer theories than the SAT to represent the models \cite{esbmc2018}.

\subsection{ESBMC }
ESBMC (or Efficient SMT-based Bounded Model Checker) is an open source, permissively licensed (Apache 2), cross platform bounded model checking for C and C++ programs \cite{esbmc2018}. 
%The efficient SMT based model checker, is a software verification tool for C and C++ code bases. 
Fig. \ref{fig:esbmcarch} shows the tool architecture.

\begin{figure}[h]
\includegraphics[width=0.4\textwidth]{ESBMCarch4}
\centering
\caption{ESBMC architecture (adapted from \cite{esbmc2018}).}
\label{fig:esbmcarch}
\end{figure}
ESBMC uses clang as front-end, a state-of-the-art compiler suite for C/C++/ObjectiveC/ObjectiveC++ widely used in industry \cite{esbmc2018}. % and generate an Abstract Syntax Tree (AST). %One legacy CBMC-based front-end that supports both C and C++ and a new clang-based front-end that currently only supports C. The data types and bit vectors are created in the front-end when parsing the code. 
%Regardless of the chosen front-end, 
The output is an AST (tree representation of the abstract syntactic structure of source code written in a programming language) that will be used by the GOTO converter to generate a GOTO program, which has simplified control flow and is suitable for bounded unwinding. The next step is the symbolic execution, when the GOTO program is executed (unrolling loops up to the bound $ k $) and converted to Static Single Assignments (SSA) form. SSA form enables the simplification of algorithms and reduction of computational complexity. %, where each variable is a target of exactly one assignment statement in the program text. 
The feature called k-induction at Fig. \ref{fig:esbmcarch} allows BMC to find a property violation or even to prove (partial) correctness without fully unwinding loops. 
%
%During the symbolic execution, ESBMC aggressively tries to simplify the program \cite{Ramalho}; it propagates all constants and solves any assertions that can be statically determined. This is an important step for the verification, since ESBMC can fully verify programs without calling a solver, if the inputs are deterministic. That characteristic is useful at this work to perform the sizing check of the PV system project. 
Following, the SSA expressions are then encoded using the chosen SMT solver supported by ESBMC. At this paper, it was used the Z3 solver \cite{DeMoura}.
%: Boolector (default) \cite{Brummayer}, Z3 \cite{DeMoura}, MathSAT \cite{Cimatti}, Yices \cite{Dutertre} and CVC4 \cite{Barrett}.

@@

%\subsection{Cyber-physical systems x Energy x ESBMC}
%According \cite{UC}, cyber-physical systems (CPS) are integrations of computation, networking, and physical processes. Embedded computers and networks monitor and control the physical processes, with feedback loops where physical processes affect computations and vice-versa. The economic and societal potential of such systems is vastly greater than what has been realized, and major investments are being made worldwide to develop the technology. The technology builds on the discipline of embedded systems (computers and software embedded in devices) whose principle mission is not computation, such as cars, toys, medical devices, and scientific instruments. CPS integrates the dynamics of the physical processes with those of the software and networking, providing abstractions and modeling, design, and analysis techniques for the integrated whole. 
%
%Energy production, distribution, and optimization are all CPS problems, mas mentioned in \cite{UC}. For example, the smart grid combines multiple electric power production plants with a multiplicity of loads using dynamic load balancing and dynamic pricing with demand-response strategies. Smart buildings integrate sensors into control systems for lighting, HVAC (heating, ventilation, and air conditioning), and safety (as fire monitoring and evacuation). 
%
%Therefore, in the context of this proposal, a cyber-physical system, specifically a solar photovoltaic system, will be mathematically modeled. Then a model checking will be performed to do a verification of the project in order to guarantee that the design of a stand-alone PV solution will meet the requirements, specifications and regulations. 
%
%As described in \cite{Alur}, in program verification, it is aimed to check if a program satisfies its logical specification. Contemporary verification tools vary widely in terms of source languages, verification methodology, and the degree of automation, but they all rely on repeatedly invoking an SMT solver. This work tracks the same path. 
%
%----------------------------------------
\section{Methodology }
\label{sec:Methodology}
%----------------------------------------
%
%Fig. \ref{fig:statetransition} shows the state diagram for the proposed tool. 
%
%\begin{figure}[h]
%\includegraphics[width=0.5\textwidth]{statetransition}
%\centering
%\caption{State transition diagram to the proposed automated verification.}
%\label{fig:statetransition}
%\end{figure}
%
%It is worth mentioning that the steps described here are carried out shortly after the design of the solar photovoltaic system, as soon as the equipment and equipment specifications are defined; before buying and deploying them, as a way to ensure that the project will not fail. 
%
The flowchart of the proposed (automated) verification method is illustrated in Fig.~\ref{fig:flowchartgeneral}. 
The PV input data (load power demand, load consumption of energy, the PV system project sizing data, and the manufacturer's data-sheet information), the formulas to check the sizing project, the mathematical model, the limits of the weather non-deterministic variables, are all written as an ANSI-C code %\cite{ANSIC}.

@@

In a nutshell, ESBMC will process the C code with constraints ($C$) and properties ($P$) from the PV system that were inputted by the user, and the tool will verify automatically if the requirements are met. If the return is a fail, then the tool details the condition of violation, i.e., the state of each variable related to the system. That information can be used as a feedback to improve the PV system project. In the other hand, if the verification succeeds, as indicated in Fig. \ref{fig:esbmcarch}, the PV system will present its intended behavior at the field.

@@

Algorithm~\ref{alg:verification-algorithm} describes the pseudo code used to perform the automated verification. The line 1 indicates function call that performs the size checking of the entire PV system: using \eqref{eq:NTPmin}, \eqref{eq:NTP}, \eqref{eq:NPSmin}, \eqref{eq:NPS}, \eqref{eq:NPP}, and \eqref{eq:NPPmin} to verify the PV panel; using \eqref{eq:Cbank}, \eqref{eq:Nbtotal}, and \eqref{eq:batcheck} to verify the batteries; using \eqref{eq:vcvsystem}, \eqref{eq:icmin}, and \eqref{eq:icicmin} to verify the charge controller; and using \eqref{eq:vindc}, \eqref{eq:voutac}, and \eqref{eq:invcheck} to verify the inverter. The verification is done by standard \textit{assert} macro from C language to each equation. The argument to \textit{assert} must be true when the macro is executed, otherwise the program aborts and prints an error message. If there is no fail, the algorithm goes on and the batteries are assumed to have SOC of 80\% (line 5).

%At both case studies, it was considered 21 years of historic data from Manaus. 
%According with X and Y, t
The annual average temperature ($T$) in Manaus is from 23$^{o}$C to 32$^{o}$C; and irradiance ($G$) varies from 274 W/m$^{2}$ to 852 W/m$^{2}$ when there is sunlight. That information is passed to the C code at lines 7 to 10 using the feature of non-deterministic variable from ESBMC to explore all the possible states from the variables and the \textit{assume} macro to limit the values considering a range. 
%and irradiance varies from 0 W/m$^{2}$ to 852 W/m$^{2}$ (with minimum of 274 W/m$^{2}$ during the daytime, when there is sunlight). 
Another characteristic of Manaus, is that the PV generation can be despised from 16:00h to the 8:00h of the next day, independent of the season.
%there is PV generation only between 8:00 h and 16:00 h every day, 
%with zero electric energy generation from 18:00 h to 6:00 h of the next day; and with insignificant generation from 6:00 h to 8:00 h, and from 16:00 h to 18:00 h of the same day. 
Therefore, in order to reduce the computational effort,
% caused by the state explosion inherent of the technique, 
the 24h-day was divided, at the algorithm, in 3 parts of 8 h each (8:00h-16:00h, 16:00h-24:00h, and 00:00h-8:00h); and only the 8:00h-16:00h period with PV generation. The time step was adjusted to 1 hour, and from line 1 to line 29 the algorithm perform the verification from 8;00h to 16:00h, where there is PV generation. Therefore, the model from PV generator is used at the function call of line 11, to produce the voltage and current considering the actual $G$ and $T$. Following, related to every hour considered, the conditional \textit{if-else-endif} statements from lines 12, 17, 23 and 27, will perform the charge and discharge of batteries according the PV generation, the updated state of charge of batteries, the house's load and the set-up information of the system.
After this process is started the battery autonomy verification, from line 30. Based on the fact that won't be PV generation after 16:00h, the algorithm will only discharge the batteries until a new charging process (at the next day) to start. The formal verification is guaranteed by \textit{assert} macro to specific variables of the model, according lines 
% and the non-deterministic variables $G$ and $T$ are considered during the formal verification of the system, otherwise, during the other two periods, there is no PV generation and just the power consumption from the backup batteries. 
%Within this 8h-period, $G$ and $T$ are automated verified with different values every one hour.
%, and change their value every 1 h according with the algorithm created using the technique.
